---
title: "ESM_206_Lab_5"
author: "Shelby Walsh"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for differnt groups
- use {lubridate} to parse dates
- find confident intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}

library(tidyverse)
library(here)
library(janitor)

```

```{r}
# import lobster data

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"), na = "-99999") %>% 
  clean_names()

```

Use tidyr::uncount() nfunction to expand a frequency variable
```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

Yay, each lobster has its own row. TIDY FORMAT!!!!

Let's look at this a bunch of ways

```{r}

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")


```

Use `lubridate` package to parse dates and times:

I'm going to add a new c olumn with `mutate` that contains data with actual date
```{r}
# Lobsterdate

lobster_date <-lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
    )

```

Parse `Date` column to get different pieces separated

```{r}

lobster_parse_date <-lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

```

Count lobsters by difffernt groupings...

Let's say I want to count # lobsters by year and month

```{r}

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site

# group_by + summarize + n()

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarise(
    mean_size = mean(size_mm, na.rm=TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )
```

Confident intervals and t-tests

Use the `t_test()` function to find confidence intervals and perform t-test

```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)

```

Two sample t.test:
Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef?

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)
mohk_napl_ttest
```

Mean Lobster size differed significantly between Mohawk and Naples Reefs (t(`r mohk_napl_ttest$parameter`)= 19.849)....

Maybe an easier way...

```{r}

Lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <-t.test(size_mm ~ site, data = Lobster_2sample)
ttest_2
```

Make a geom_tile Heatmap

```{r}

ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

